When does vapor pressure deficit drive or reduce evapotranspiration?

### 2017-10-04 doc/paper/vpd_et_paper.tex
finished with spell check
minor formatting changes

\section{Introduction}\explain{This section needs to be fleshed out, and I definately need to add more citations}

Changes to vapor pressure deficit (VPD) alter the atmospheric demand for water from the land surface. However, plant stomata have evolved to optimally regulate the exchange of water and carbon between vegetation and the atmosphere \citep{Franks_2017}. Therfore, an increase (decrease) in VPD may not correspond to an increase (decrease) in evapotranspiration (ET) because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.

Quantifying the plant response to a perturbation to atmospheric VPD increases our understanidng of feedbacks between the land surface and the atmosphere. If plant reponse reduces ET in response to an increase in VPD, the land surface will contribute a positive feedback in reponse to atmospheric drying. Conversely, if plant response increases ET in response to increase in VPD, then the land surface will contribute a negative feedback to atmospheric drying. The sign of these feedbacks drives the evolution of the atmosphere and landsurface at many timescales, from diurnal to interdecadal.

Here we use a Penman-Monteith framework to quantify plant reponse to perturbations to atmospheric demand for water. Section 2 derives the framework, Section 3 describes the data used, Section 4 presents results, and Section 5 discusses conclusions. The goal of this paper is to use reasonable approximations as a tool to increase intuition for plant response to atmospheric drying. This intuition will aid interpretation of observations and full complexity climate models.

@@

The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variaibles:

@@

 where variable definitions are given in Table 1. \cite{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empiracle approaches. The result for leaf-scale stomatal resistance was:

@@

While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a relatively strong function of ET. To remove dependence of ET on $A$ we can use the semi-impiracle results of \cite{Zhou_2015}. \cite{Zhou_2015} showed that:

@@

We can then take the derivative with respect to $D$ to determine ecosystem reponse to atmospheric demand perturbations:

@@

This ``pseudo-LAI'' is some part ``true'' LAI (a measure of leaf area), and some part model and observational error, including error involving our assumption of constant uWUE. By calculating a unique LAI for each observation we will progate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.

@@

We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th pecentile). Also, because some sites use half hourly data but some use hourly, we aggregate all data to hourly averages. Only times with good quality control flags are used.

@@

By construction, the variability in the LAI term (Equation \ref{lai}) contains all model and observational uncertities. LAI also has physical meaning corresponding to ``true'' leaf area, and we expect that it would be approximately $O(1)$. We can have some confidence in our framework, including the assumption of constant uWUE, if calculated LAIs are generally $O(1)$. Figure \ref{lai_fig} presents the histogram of calculated LAIs with outliers (lowest and highest 5\% percent) and unphysical values (LAI $<$ 0.) removed. All remaining LAI values are $O(1)$ which provides confidence in model framework.

@@

An additional concern is that the LAI term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative. Figure \ref{lai_vpd_fig} plots the joint distrivution of LAI and VPD, and shows that LAI is very weakly a function of VPD. Given this weak dependence, we argue that Equation \ref{d_et} is a valid approxiamtion for ET response to $D$.

@@

In Term 1, $\frac{P}{T} \propto \rho$, so this should vary little relative to $g_a$ and $\Delta$.  $\gamma$ should also be relatively constant, So the scaling term, Term 1, should be primarily a function of $g_a$ and temperature (through the function $\Delta$). While temperature range may vary for PFT, the fucntional form of $\Delta$ will be the same. $g_a$ will vary strongly with PFT due to the importance of surface roughness. However, the coeffiecent of variability for $g_a$ is relatively constant across PFT. So, the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similiar across PFT.

Figure \ref{scale_vary}A shows Term 1 normalized by mean $g_a$ (calcualted for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the $g_a$ term's relative variability. Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$. % plot every PFT and show they collapse onto the same vurve?

While the relative variability of Term 1 is similiar across PFT, the absolute value of Term 1 varies strongly across PFT. Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT. As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger. Systematic differences in observed temperatures also cause differences in the average magnitude of Term 1. For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term. Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity. To summarize, the variability of Term 1 will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to Figure \ref{scale_vary}B.

@@

Term 2 minus Term 3 determines the sign and magnitude that the scaling Term 1 is multiplied by. If we assume that $c_s$ variability is relatively less than LAI and $D$ variability, then variability within PFT will be soley detmined by LAI and $D$. Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and LAI, as a function of PFT. In Figure \ref{term3}a lower uWUE and LAI shift the disrtribution of (Term 2 - Term 3) towards negative values. Additionally, the smaller $g1$, the greater the relative $D$ dependence of of (Term 2 - Term 3). This is observed most strongly for the ENF PFT, which has the smallest $g1$ (74.31).

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{./fig05.pdf}
\caption{Sources of variability for Term 2 - Term 3. Top: Term 2 - Term 3 as a function of VPD, with LAI held fixed at PFT averages. The observed range of VPD for each PFT is also shown below the x-axis. Line extent correspondes to 5th and 95th precentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.

@@

Table 3 confirms these expectations for PFT behaviour of $\frac{\partial \; ET}{\partial \; D}$. For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; D}$ is less than zero. However, $\frac{\partial \; ET}{\partial \; D}$ evaluated at the average of all variables (e.g. LAI, $T$, $c_s$, $D$) is only negative for CSH and GRA. And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; D}$ < 0 less than half the time. These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
n
The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability. To understand $D$'s contribution better, we use a linear approximation and present $\frac{\partial \; ET}{\partial \; D}$ multiplied by $D$'s standard deviation. The range of $D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$. Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiation generally the driver of ET (cite joe berry here). For all PFTs except for CSH $D$ contributes between 14.5 - 20.5 \% of $R$'s contribution to variability. For CSH the portion is much larger, about 44 \%. However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variabilility, so these values are just estimates of $D$'s contribution to ET's variability.

So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; D}$ and how it varies with PFT. Large mean LAI and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$. However, the strongly nonlinear funtion of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$). ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; D}$ on $D$, and makes the function more strongly nonlinear. This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative futher than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$. GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this). Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}). Variability in $D$ accounts for the largest about of $ET$ variability for CSH. For the other PFTs, $D$ contributes less to $ET$ variaibiliyt, but still represents about 15-20 \% of $R$'s contributions to ET variability.

@@

Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to intepret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT. Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary. Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, and each row is a different PFT.

The full observations generally confirm expectations from Section \ref{results}. CRO has the most postitive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and reponse depends more with the environmental conditions for the other PFTs (especially ENF). Through the columns of Figure \ref{real} we can see the impact of LAI and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$. $g_a$'s (columns 1 and 3) scaling alters the magnitude considerably. $LAI$ (columns 2 and 3) variability adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given LAI's role in representing model and observational uncertainty. However, the general story with the noise appears to match the cleaner signal when $LAI$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these pltos better - way too much overlapping of points that hurts the story}. One exception is possibly with GRA, for which uncertainty is high and causes the full complexity plots (Columns 1 and 2) to not match well with LAI held fixed (Columns 3 and 4).

For ENF and GRA $D_{ETmin}$ does not appear to be only a function of LAI. It turns out that the site to site variaibility in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the pregious section. The impact is observable in both ENF and GRA, but especially for ENF which has a larger $\frac{\partial^2 \; ET}{\partial^2 \; D}$ than the other PFTs.

In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivty to LAI measures of uncertainty observed in Figure \ref{term3}. Our LAI-based method of uncertainty propogation blurrs the idealized expectations the most for GRA, and also has a considerable effect for CRO. We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when LAI is held fixed. **see somewhat preferred alternate figure \ref{real2} \explain{I think I like this more as thinking in terms of T and RH is easier. However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another diminsion). I could just include both versions of figure 6 though (using figure 6a as a bridge to figure 6b)}.
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{./fig06.png}
\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; D}$. Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labelled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with LAI held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with LAI held fixed. For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn. Please note differences in the colorbar scale.}
\label{real}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{./fig06b.png}
\caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; D}$. Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labelled. If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{LAI}$. }
\label{real2}
\end{figure}

@@

The idealized framework leads to the following general conclusions:
\begin{itemize}
  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; D}$. This is a leading order effect for observing higher magnitude reponses in DBF and ENF.
\item In general, CSH has the most negative (i.e. ET reduced) response to increases in $D$ (atmospheric drying). So CSH plants will almost always try and conserve water, effectivey reducing ET with dry atmospheric perturbation.
\item CRO has the most postive response (i.e. ET increased) in response to increases in $D$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
\item The reponse is more a function of the environment for DBF, ENF, and GRA. Because as VPD increases the response is more likely to be postive, if RH is fixed then the reponse will be more likely to be positive at warmer T, or if T is fixed the response is more likely to be positive with decreasing RH.
\item ENF has the strongest dependence on environmental conditions due to its small $g1$.
\item Model and observational uncertainty is highest for GRA and CRO, so conlusions about those PFTs should be tempered.
\item However, inclusion of uncertainty doesn't alter conclusions about DBF, ENF, and CSH.
\end{itemize}

>>>>

\section{Introduction}\explain{This section needs to be fleshed out, and I definitely need to add more citations}

Changes to vapor pressure deficit (VPD) alter the atmospheric demand for water from the land surface. However, plant stomata have evolved to optimally regulate the exchange of water and carbon between vegetation and the atmosphere \citep{Franks_2017}. Therefore, an increase (decrease) in VPD may not correspond to an increase (decrease) in evapotranspiration (ET) because stomatal closure (opening) can cancel the effects of shifts to atmospheric demand.

Quantifying the plant response to a perturbation to atmospheric VPD increases our understanding of feedbacks between the land surface and the atmosphere. If plant response reduces ET in response to an increase in VPD, the land surface will contribute a positive feedback in response to atmospheric drying. Conversely, if plant response increases ET in response to increase in VPD, then the land surface will contribute a negative feedback to atmospheric drying. The sign of these feedbacks drives the evolution of the atmosphere and land-surface at many timescales, from diurnal to interdecadal.

Here we use a Penman-Monteith framework to quantify plant response to perturbations to atmospheric demand for water. Section 2 derives the framework, Section 3 describes the data used, Section 4 presents results, and Section 5 discusses conclusions. The goal of this paper is to use reasonable approximations as a tool to increase intuition for plant response to atmospheric drying. This intuition will aid interpretation of observations and full complexity climate models.

@@

The Penman-Monteith equation (hereafter PM) estimates ET as a function of atmospheric and land-surface variables:

@@

 where variable definitions are given in Table 1. \cite{MEDLYN_2011} developed a model for $g_s$ by combining optimal photosynthesis theory with empirical approaches. The result for leaf-scale stomatal resistance was:

@@

While Equation 3 can be used in PM, it will make analytical work with the function intractable because $A$ is a relatively strong function of ET. To remove dependence of ET on $A$ we can use the semi-empirical results of \cite{Zhou_2015}. \cite{Zhou_2015} showed that:

@@

We can then take the derivative with respect to $D$ to determine ecosystem response to atmospheric demand perturbations:

@@

This ``pseudo-LAI'' is some part ``true'' LAI (a measure of leaf area), and some part model and observational error, including error involving our assumption of constant uWUE. By calculating a unique LAI for each observation we will propagate any model and observational uncertainty forward into our expression for $\frac{\partial \; ET}{\partial D}$.

@@

We restrict our analysis to the daytime (sensible heat > 5 W m$^{-1}$ and shortwave radiation > 50 W m$^{-2}$) when there is no precipitation and the plants are growing (GPP > 10\% of the 95th percentile). Also, because some sites use half hourly data but some use hourly, we aggregate all data to hourly averages. Only times with good quality control flags are used.

@@

By construction, the variability in the LAI term (Equation \ref{lai}) contains all model and observational uncertainties. LAI also has physical meaning corresponding to ``true'' leaf area, and we expect that it would be approximately $O(1)$. We can have some confidence in our framework, including the assumption of constant uWUE, if calculated LAIs are generally $O(1)$. Figure \ref{lai_fig} presents the histogram of calculated LAIs with outliers (lowest and highest 5\% percent) and nonphysical values (LAI $<$ 0.) removed. All remaining LAI values are $O(1)$ which provides confidence in model framework.

@@

An additional concern is that the LAI term may in fact be some function of $D$, in which case the dependence would need to be accounted for when taking the derivative. Figure \ref{lai_vpd_fig} plots the joint distribution of LAI and VPD, and shows that LAI is very weakly a function of VPD. Given this weak dependence, we argue that Equation \ref{d_et} is a valid approximation for ET response to $D$.

@@

In Term 1, $\frac{P}{T} \propto \rho$, so this should vary little relative to $g_a$ and $\Delta$.  $\gamma$ should also be relatively constant, So the scaling term, Term 1, should be primarily a function of $g_a$ and temperature (through the function $\Delta$). While temperature range may vary for PFT, the functional form of $\Delta$ will be the same. $g_a$ will vary strongly with PFT due to the importance of surface roughness. However, the coefficient of variability for $g_a$ is relatively constant across PFT. So, the influence of $g_a$ on the relative (to the mean) variability of Term 1 is approximately similar across PFT.

Figure \ref{scale_vary}A shows Term 1 normalized by mean $g_a$ (calculated for each plant functional type), and confirms that much of the relative variability of Term 1 is contained in the $g_a$ term's relative variability. Additionally, the impact of $T$ on the relative variability increases with increasing $g_a$. % plot every PFT and show they collapse onto the same vurve?

While the relative variability of Term 1 is similar across PFT, the absolute value of Term 1 varies strongly across PFT. Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT. As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger. Systematic differences in observed temperatures also cause differences in the average magnitude of Term 1. For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term. Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity. To summarize, the variability of Term 1 will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to Figure \ref{scale_vary}B.

@@

Term 2 minus Term 3 determines the sign and magnitude that the scaling Term 1 is multiplied by. If we assume that $c_s$ variability is relatively less than LAI and $D$ variability, then variability within PFT will be solely determined by LAI and $D$. Figure \ref{term3} shows how (Term 2 - Term 3) varies with $D$ and LAI, as a function of PFT. In Figure \ref{term3}a lower uWUE and LAI shift the distribution of (Term 2 - Term 3) towards negative values. Additionally, the smaller $g1$, the greater the relative $D$ dependence of of (Term 2 - Term 3). This is observed most strongly for the ENF PFT, which has the smallest $g1$ (74.31).

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{./fig05.pdf}
\caption{Sources of variability for Term 2 - Term 3. Top: Term 2 - Term 3 as a function of VPD, with LAI held fixed at PFT averages. The observed range of VPD for each PFT is also shown below the x-axis. Line extent corresponds to 5th and 95th percentiles, while stars denote the location of the 25th, 50th, and 75th percentiles.

@@

Table 3 confirms these expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; D}$. For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; D}$ is less than zero. However, $\frac{\partial \; ET}{\partial \; D}$ evaluated at the average of all variables (e.g. LAI, $T$, $c_s$, $D$) is only negative for CSH and GRA. And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; D}$ < 0 less than half the time. These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.
n
The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability. To understand $D$'s contribution better, we use a linear approximation and present $\frac{\partial \; ET}{\partial \; D}$ multiplied by $D$'s standard deviation. The range of $D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$. Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiation generally the driver of ET (cite joe berry here). For all PFTs except for CSH $D$ contributes between 14.5 - 20.5 \% of $R$'s contribution to variability. For CSH the portion is much larger, about 44 \%. However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variability, so these values are just estimates of $D$'s contribution to ET's variability.

So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; D}$ and how it varies with PFT. Large mean LAI and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$. However, the strongly nonlinear function of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$). ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; D}$ on $D$, and makes the function more strongly nonlinear. This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative further than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$. GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magnitude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this). Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}). Variability in $D$ accounts for the largest about of $ET$ variability for CSH. For the other PFTs, $D$ contributes less to $ET$ variability, but still represents about 15-20 \% of $R$'s contributions to ET variability.

@@

Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to interpret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT. Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary. Each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, and each row is a different PFT.

The full observations generally confirm expectations from Section \ref{results}. CRO has the most positive values of $\frac{\partial \; ET}{\partial \; D}$, $\frac{\partial \; ET}{\partial \; D}$ is almost always negative for CSH, and response depends more with the environmental conditions for the other PFTs (especially ENF). Through the columns of Figure \ref{real} we can see the impact of LAI and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; D}$. $g_a$'s (columns 1 and 3) scaling alters the magnitude considerably. $LAI$ (columns 2 and 3) variability adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; D}$, which is slightly undesirable given LAI's role in representing model and observational uncertainty. However, the general story with the noise appears to match the cleaner signal when $LAI$ is help constant and $D_{ETmin}$ is clearly visible \explain{I really need to make these plots better - way too much overlapping of points that hurts the story}. One exception is possibly with GRA, for which uncertainty is high and causes the full complexity plots (Columns 1 and 2) to not match well with LAI held fixed (Columns 3 and 4).

For ENF and GRA $D_{ETmin}$ does not appear to be only a function of LAI. It turns out that the site to site variability in $\gamma$ causes $D_{ETmin}$ to vary, which is not discussed in the previous section. The impact is observable in both ENF and GRA, but especially for ENF which has a larger $\frac{\partial^2 \; ET}{\partial^2 \; D}$ than the other PFTs.

In general the full complexity plots of $\frac{\partial \; ET}{\partial \; D}$ match our expectations, even with the large sensitivity to LAI measures of uncertainty observed in Figure \ref{term3}. Our LAI-based method of uncertainty propagation blurs the idealized expectations the most for GRA, and also has a considerable effect for CRO. We therefor have the most confidence in our conclusion based on Equation \ref{d_et} for PFTS CSH, DBF, and ENF, as the full complexity plots with uncertainty included closely match the story when LAI is held fixed. **see somewhat preferred alternate figure \ref{real2} \explain{I think I like this more as thinking in terms of T and RH is easier. However, I used the other plot because Fig 5 does not discuss things in terms of temperature, as this would make things more complicated (adding another dimension). I could just include both versions of figure 6 though (using figure 6a as a bridge to figure 6b)}.
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{./fig06.png}
\caption{Scatter plots of $\frac{\partial \; ET}{\partial \; D}$. Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled: Column 1 - $\frac{\partial \; ET}{\partial \; D}$; Column 2 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$; Column 3 - $\frac{\partial \; ET}{\partial \; D}$ with LAI held fixed at PFT average; and Column 4 - $\frac{\partial \; ET}{\partial \; D}$ normalized by $g_a$ and with LAI held fixed. For reference, lines corresponding to RH = 20\% and RH = 90 \% are drawn. Please note differences in the colorbar scale.}
\label{real}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{./fig06b.png}
\caption{****alternate Fig 06****  Scatter plots of $\frac{\partial \; ET}{\partial \; D}$. Each row is a different PFT, and each column is a different quantity related to $\frac{\partial \; ET}{\partial \; D}$, as labeled. If I end up using this, I could also draw on the curve of $D_{ETmin}$ with $\overline{LAI}$. }
\label{real2}
\end{figure}

@@

The idealized framework leads to the following general conclusions:
\begin{itemize}
  \item Aerodynamic resistance plays an important role of scaling $\frac{\partial \; ET}{\partial \; D}$. This is a leading order effect for observing higher magnitude responses in DBF and ENF.
\item In general, CSH has the most negative (i.e. ET reduced) response to increases in $D$ (atmospheric drying). So CSH plants will almost always try and conserve water, effectively reducing ET with dry atmospheric perturbation.
\item CRO has the most positive response (i.e. ET increased) in response to increases in $D$. This is consistent with CROs that may be evolved or bred to thrive in non-water-limited environments.
\item The response is more a function of the environment for DBF, ENF, and GRA. Because as VPD increases the response is more likely to be positive, if RH is fixed then the response will be more likely to be positive at warmer T, or if T is fixed the response is more likely to be positive with decreasing RH.
\item ENF has the strongest dependence on environmental conditions due to its small $g1$.
\item Model and observational uncertainty is highest for GRA and CRO, so conclusions about those PFTs should be tempered.
\item However, inclusion of uncertainty doesn't alter conclusions about DBF, ENF, and CSH.
\end{itemize}
